---
title: "p8105_hw3_sc4935"
author: "Shiwei Chen"
date: "10/13/2021"
output: github_document

---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggridges)
library(knitr)
```


# Problem 1

Load dataset of instacart.

```{r}
library(p8105.datasets)
data("instacart")
```

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart_df = count(instacart, aisle) %>% 
  mutate(aisle_ranking = min_rank(desc(n)))
instacart_df 

nrow(instacart_df)
arrange(instacart_df, aisle_ranking) %>% 
  filter(aisle_ranking == 1)
view(instacart_df)
```

So, there are 134 kinds of aisles, and by arranging them, the most items ordered from fresh vegetables.


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart_df %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = n, y = aisle)) +
  geom_point() + 
  labs(
    title = "The plot of items in each aisle",
    x = "Number of items ordered in each aisle",
    y = "Aisle name",
    caption = "Data from instacart of the p8105.datasets package"
  )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart_table1_df = instacart %>% 
  select(aisle, product_name) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(order_times = n()) %>% 
  mutate(product_rank = min_rank(desc(order_times))) %>% 
  filter(product_rank == 1) %>% 
  select(-product_rank) %>% 
  pivot_wider(
  names_from = "product_name", 
  values_from = "order_times") %>% 
  view()

knitr::kable(instacart_table1_df)
instacart_table1_df 
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart_table2_df = instacart %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(order_dow, product_name) %>% 
  summarise(mean_hours = mean(order_hour_of_day)) %>% 
  mutate(order_dow = recode(order_dow, "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", "4" = "Thursday", "5" = "Friday", "6" = "Saturday")) %>% 
  pivot_wider(
  names_from = "order_dow", 
  values_from = "mean_hours") %>% 
  view()

knitr::kable(instacart_table2_df)
instacart_table2_df
```

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.











# Problem 2

Load dataset of BRFSS (brfss_smart2010).

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

First, do some data cleaning.

format the data to use appropriate variable names.
focus on the “Overall Health” topic.
include only responses from “Excellent” to “Poor”.
organize responses as a factor taking levels ordered from “Poor” to “Excellent”.

```{r}
brfss_cleaned_df = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  select(-locationabbr) %>% 
  separate(locationdesc, into = c("state", "place"), sep = " - ") %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  view()

brfss_cleaned_df  
```

In 2002, which states were observed at 7 or more locations? 

```{r}
brfss_2002 = filter(brfss_cleaned_df, year == 2002) %>% 
  group_by(state) %>% 
  summarize(observed = n_distinct(place)) %>% 
  filter(observed >= 7) %>% 
  view()
brfss_2002
```

What about in 2010?

```{r}
brfss_2010 = filter(brfss_cleaned_df, year == 2010) %>% 
  group_by(state) %>% 
  summarize(observed = n_distinct(place)) %>% 
  filter(observed >= 7) %>% 
  view()
brfss_2010
```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent_df = filter(brfss_cleaned_df, response == "Excellent") %>% 
  select(year, state, place, data_value) %>% 
  group_by(year, state) %>% 
  summarise(mean_dv = mean(data_value)) %>% 
  view()

excellent_df
ggplot(excellent_df, aes(x = year, y = mean_dv, color = state)) +
  geom_point(alpha = 0.5) + geom_line(alpha = 0.5) + 
  labs(
    title = "The average value across locations over time within a state",
    x = "Average of the data_value",
    y = "Year",
    caption = "Data from brfss_smart2010 of the p8105.datasets package"
  ) 
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
two_panel_plot = brfss_cleaned_df %>% 
  filter(state == "NY") %>% 
  filter(year == 2006 | year == 2010) %>% 
  select(year, state, place, response, data_value) %>% 
  view()

two_panel_plot

ggplot(two_panel_plot, aes(x = data_value, y = response, color = place)) + 
  geom_point(alpha = 1) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ year) +
  labs(
    title = "Distribution of data_value for responses among locations in NY for 2006 and 2010",
    x = "Data_value",
    y = "Responses",
    caption = "Data from brfss_smart2010 of the p8105.datasets package"
  ) 
```






# Problem 3

Load dataset of accel_data.

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  select(-day_id) %>% 
  mutate(day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday",  "Sunday"))) %>% 
  arrange(week, day) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "counts", 
    values_to = "activity") %>% 
  mutate(weekday_vs_weekend = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>%
  separate(counts, into = c("remove", "minutes"), sep = "_") %>% 
  select(-remove) %>% 
  mutate(week = as.character(week)) %>% 
  mutate(minutes = as.numeric(minutes)) %>% 
  view()
  
accel_df  
```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_df %>% 
  group_by(week, day) %>% 
  summarise(total_activity = sum(activity)) %>% 
   pivot_wider(
  names_from = "day", 
  values_from = "total_activity") %>% 
  view() %>% 
  knitr::kable()
```

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_df %>% 
  ggplot(aes(x = minutes, y = activity, color = day)) + 
  geom_point(alpha = 0.5) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Minutes of a day",
    y = "Activity count",
    caption = "Data from accel_data.csv"
  ) 
  

```