---
title: "p8105_hw3_sc4935"
author: "Shiwei Chen"
date: "10/13/2021"
output: github_document

---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggridges)
library(knitr)
```


# Problem 1

Load dataset of instacart.

```{r}
library(p8105.datasets)
data("instacart")
```

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart_df = count(instacart, aisle) %>% 
  mutate(aisle_ranking = min_rank(desc(n)))
instacart_df 

nrow(instacart_df)
arrange(instacart_df, aisle_ranking) %>% 
  filter(aisle_ranking == 1)
view(instacart_df)
```

So, there are 134 kinds of aisles, and by arranging them, the most items ordered from fresh vegetables.


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart_df %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = n, y = aisle)) +
  geom_point() + 
  labs(
    title = "The plot of items in each aisle",
    x = "Number of items ordered in each aisle",
    y = "Aisle name",
    caption = "Data from instacart of the p8105.datasets package"
  )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart_table1_df = instacart %>% 
  select(aisle, product_name) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(order_times = n()) %>% 
  mutate(product_rank = min_rank(desc(order_times))) %>% 
  filter(product_rank == 1) %>% 
  select(-product_rank) %>% 
  pivot_wider(
  names_from = "product_name", 
  values_from = "order_times") %>% 
  view()

instacart_table1_df 
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart_table2_df = instacart %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(order_dow, product_name) %>% 
  summarise(mean_hours = mean(order_hour_of_day)) %>% 
  mutate(order_dow = recode(order_dow, "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", "4" = "Thursday", "5" = "Friday", "6" = "Saturday")) %>% 
  pivot_wider(
  names_from = "order_dow", 
  values_from = "mean_hours") %>% 
  view()

knitr::kable(instacart_table2_df)
instacart_table2_df
```

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

